A Memory-Bounded, Deterministic and Terminating Semantics for the Synchronous Programming Language~\CEU

### 2018-02-23 03-semantics.tex
Sugestões do Hermann
(Faltou definicao de E)
Comprime os o conteúdo da Seção 3.2

\newcommand{\NST}{\1\xrightarrow[\mathit{nst}]\1}
\newcommand{\OUT}{\1\xrightarrow[\mathit{out}]\1}
\newcommand{\LL}{\langle}
\newcommand{\RR}{\rangle}
\newcommand{\DS}{\displaystyle}

\newcommand{\1}{\;}
\newcommand{\2}{\;\;}
\newcommand{\3}{\;\;\;}
\newcommand{\5}{\;\;\;\;\;}

@@

The grammar below defines the syntax of a subset of \CEU that is
sufficient to describe all semantic peculiarities of the language.
\bgroup
\def\lbl#1{\quad\text{\emph{#1}}}%
\newdimen\X
\X=-1\jot
\begin{alignat*}{2}
  p\Coloneqq
      &\enspace\ceu{\Mem(\Id)}
      &&\lbl{any memory access to~``$\ceu\Id$''}\\[\X]
      %%
  \mid&\enspace\ceu{\AwaitExt(\Id)}
      &&\lbl{await external event~``$\ceu\Id$''}\\[\X]
      %%
  \mid&\enspace\ceu{\AwaitInt(\Id)}
      &&\lbl{await internal event~``$\ceu\Id$''}\\[\X]
      %%
  \mid&\enspace\ceu{\EmitInt(\Id)}
      &&\lbl{emit internal event~``$\ceu\Id$''}\\[\X]
      %%
  \mid&\enspace\ceu{\Break}
      &&\lbl{loop escape}\\[\X]
      %%
  \mid&\enspace\ceu{\IfElse{\Mem(\Id)}{p_1}{p_2}}
      &&\lbl{conditional}\\[\X]
      %%
  \mid&\enspace\ceu{p_1\,;\,p_2}
      &&\lbl{sequence}\\[\X]
      %%
  \mid&\enspace\ceu{\Loop p_1}
      &&\lbl{repetition}\\[\X]
      %%
  \mid&\enspace\ceu{\Every{\Id}\ p_1}
      &&\lbl{event iteration}\\[\X]
      %%
  \mid&\enspace\ceu{p_1\And p_2}
      &&\lbl{par/and}\\[\X]
      %%
  \mid&\enspace\ceu{p_2\Or p_2}
      &&\lbl{par/or}\\[\X]
      %%
  \mid&\enspace\ceu{\Fin p}
      &&\lbl{finalization}\\[\X]
      %%
  \mid&\enspace\ceu{p_1\AtLoop p_2}
      &&\lbl{unwinded loop}\\[\X]
      %%
  \mid&\enspace\ceu{p_1\AtAnd\ p_2}
      &&\lbl{unwinded par/and}\\[\X]
      %%
  \mid&\enspace\ceu{p_1\AtOr\ p_2}
      &&\lbl{unwinded par/or}\\[\X]
      %%
  \mid&\enspace\ceu{\CanRun(n)}
      &&\lbl{can run on stack level~$n$}\\[\X]
      %%
  \mid&\enspace\ceu{\Nop}
      &&\lbl{terminated program}
\end{alignat*}
\egroup

@@

Most expressions in the abstract language are mapped to their counterparts in
the concrete language.
The exceptions are the finalization block~$\ceu{\Fin{p}}$ and the
\texttt{@}-expressions which result from expansions in the transition rules to
be presented.

@@

The core of our semantics describes how a program reacts to a single external
input event, i.e., starting from an input event, how the program behaves and
becomes idle again to proceed to a subsequent reaction.
%
We use a set of small-step operational rules, which are designed in such a way
that at most one transition is possible at any time, resulting in deterministic
reactions.
%
The transition rules map a triple with a program~$p$, a stack level~$n$, and an
emitted event~$e$ to a modified triple as follows:
\[
  \<p,n,e>\trans\<p',n',e'>\,,
\]
where~$p,p'\in\P$ are abstract-language programs, $n,n'\in\N$ are non-negative
integers representing the current stack level, and~$e,e'\in\E\cup\{\nil\}$ are
the events emitted before and after the transition (both being possibly the
empty event~$\nil$).

@@

The~$\out$ rules \R{push} and \R{pop} are non-recursive definitions which only
apply to the program as a whole, and are the only to manipulate the stack
level:
\begin{align*}
  &\AxiomC{$e\ne\nil$}
  \UnaryInfC{$\<p,n,e>\out\<\bcast(p,e),n+1,\nil>$}
  \DisplayProof
  \Rtag{push}\\[2\jot]
  %%
  &\AxiomC{$n>0$}
  \AxiomC{$\ceu{p=\Nop}\vee\isblocked(p,n)$}
  \BinaryInfC{$\<p,n,\nil>\out\<p,n-1,\nil>$}
  \DisplayProof
  \Rtag{pop}
\end{align*}

@@

The~$\nst$ rules are recursive definitions with the following general format:
\[
\<p,n,\nil>\nst\<p',n,e>.
\]
%
%-
% \begin{align*}
% \LL p, n,\epsilon \RR &\NST
% \LL p',n,e        \RR
%     & \textbf{(rule-inner)}
% \end{align*}
%-
%
Nested transitions do not affect the stack level and never have an emitted
event as a precondition.  The distinction between~$\out$ and~$\nst$ prevents
rules \R{push} and \R{pop} from matching and, consequently, from
inadvertently modifying the current stack level before the nested reaction
is complete.

A complete reaction consists of a series of transitions:
\begin{align*}
  \<p,0,e_\ext>\outpush\<p_1,1,\nil>
  \Big[\null\nst[*]\null\out\null\Big]\!\!\ast
  \null\nst[*]\null\outpop\<p',0,\nil>\,.
\end{align*}
%
%-
% \begin{align*}
% a) &\5\5
%     \LL p,0,ext \RR
%         \1\xrightarrow[out]{push}\1
%     \LL q,1,\epsilon \RR
% \\
% b) &\5\5 \1[ \1\xrightarrow[in]{*}\1
%     \LL r,i,e \RR
%         \1\xrightarrow[out]\1
%     \LL s,j,\epsilon \RR \1]*
% \\
% c) &\5\5 \1\xrightarrow[in]{*}\1
%     \LL t,k,\epsilon \RR
%         \1\xrightarrow[out]{pop}\1
%     \LL u,0,\epsilon \RR
% \end{align*}
%-
%
First, a~$\outpush$ starts a nested reaction at level~1.
Then, a series of alternations between zero or more~$\nst$ transitions (nested reactions) and a
single~$\out$ transition (stack operation) takes place.
Finally, a last~$\outpop$ transition decrements the
stack level to~0 and terminates the reaction.

The~$\nst$ transition rules for atomic expressions are defined as follows:
\begin{align*}
  \<\ceu{\Mem(\Id)},n,\nil>
  &\nst\<\ceu{\Nop},n,\nil>\Rtag{mem}\\
  %%
  \<\ceu{\EmitInt(\Id)},n,\nil>
  &\nst\<\ceu{\CanRun(n)},n,\ceu{\Id}>\Rtag{emit-int}\\
  %%
  \<\ceu{\CanRun(n)},n,\nil>
  &\nst\<\ceu{\Nop},n,\nil>\Rtag{can-run}
\end{align*}

@@

Proceeding to compound expressions, the rules for conditionals and sequences
are straightforward:
%
\begin{gather*}
  \AxiomC{$\eval(\ceu{\Mem(\Id)})$}
  \UnaryInfC{$\<\ceu{\IfElse{\Mem(\Id)}{p}{q}},n,\nil>\nst\<p,n,\nil>$}
  \DisplayProof
  \Rtag{if-true}\\[2\jot]
  %%
  \AxiomC{$\lnot\eval(\ceu{\Mem(\Id)})$}
  \UnaryInfC{$\<\ceu{\IfElse{\Mem(\Id)}{p}{q}},n,\nil>\nst\<q,n,\nil>$}
  \DisplayProof
  \Rtag{if-false}\\[2\jot]
  %%
  \AxiomC{$\<p,n,\nil>\nst\<p',n,e>$}
  \UnaryInfC{$\<\ceu{p\,;\,q},n,\nil>\nst\<\ceu{p';\,q},n,e>$}
  \DisplayProof
  \Rtag{seq-adv}
\end{gather*}
\vskip-\belowdisplayskip
\vskip-\abovedisplayskip
\vskip2\jot
\begin{align*}
  \<\ceu{\Nop;\,q},n,\nil>&\nst\<q,n,\nil>\Rtag{seq-nop}\\[2\jot]
  %%
  \<\ceu{\Break;\,q},n,\nil>&\nst\<\ceu{\Break},n,\nil>\Rtag{seq-brk}
\end{align*}

@@

The rules for loops are analogous to sequences, but use ``\code{@}'' as
separators to properly bind breaks to their enclosing loops:
\begin{align*}
  \<\ceu{\Loop{p}},n,\nil>
  &\nst\<\ceu{p\AtLoop{p}},n,\nil>\Rtag{loop-expd}\\[2\jot]
  %%
  &\hskip-6.35em
  \AxiomC{$\<q,n,\nil>\nst\<q',n,e>$}
  \UnaryInfC{$\<\ceu{q\AtLoop{p}},n,\nil>\nst\<\ceu{q'\AtLoop{p}},n,e>$}
  \DisplayProof
  \Rtag{loop-adv}\\[2\jot]
  %%
  \<\ceu{\Nop\AtLoop{p}},n,\nil>
  &\nst\<\ceu{\Loop{p}},n,\nil>\Rtag{loop-nop}\\[2\jot]
  %%
  \<\ceu{\Break\AtLoop{p}},n,\nil>
  &\nst\<\ceu{\Nop},n,\nil>\Rtag{loop-brk}
\end{align*}

@@

The semantic rules for~$\ceu{\And}$ and~$\ceu{\Or}$ compositions
force transitions on their left branches~$p$ to occur before their right
branches~$q$:
\begin{gather*}
  \hskip-.9em
  \<\ceu{p\And{q}},n,\nil>
  \nst\<\ceu{p\AtAnd(\CanRun(n);q)},n,\nil>
  \Rtag{and-expd}\\[2\jot]
  %%
  \AxiomC{$\<p,n,\nil>\nst\<p',n,e>$}
  \UnaryInfC{$\<\ceu{p\AtAnd{q}},n,\nil>\nst\<\ceu{p'\AtAnd{q}},n,e>$}
  \DisplayProof
  \Rtag{and-adv1}\\[2\jot]
  %%
  \AxiomC{$\isblocked(p,n)$}
  \AxiomC{$\<q,n,\nil>\nst\<q',n,e>$}
  \BinaryInfC{$\<\ceu{p\AtAnd{q}},n,\nil>\nst\<\ceu{p\AtAnd{q'}},n,e>$}
  \DisplayProof
  \Rtag{and-adv2}\\[2\jot]
  %%
  \<\ceu{p\Or{q}},n,\nil>
  \nst\<\ceu{p\AtOr(\CanRun(n);q)},n,\nil>
  \Rtag{or-expd}\\[2\jot]
  %%
  \AxiomC{$\<p,n,\nil>\nst\<p',n,e>$}
  \UnaryInfC{$\<\ceu{p\AtOr{q}},n,\nil>\nst\<\ceu{p'\AtOr{q}},n,e>$}
  \DisplayProof
  \Rtag{or-adv1}\\[2\jot]
  %%
  \AxiomC{$\isblocked(p,n)$}
  \AxiomC{$\<q,n,\nil>\nst\<q',n,e>$}
  \BinaryInfC{$\<\ceu{p\AtOr{q}},n,\nil>\nst\<\ceu{p\AtOr{q'}},n,e>$}
  \DisplayProof
  \Rtag{or-adv2}
\end{gather*}

@@

In a parallel~$\ceu{\AtAnd}$, if one of the sides terminates, the composition is
simply substituted by the other side (rules \R{and-nop1} and
\R{and-nop2} below).
%
In a parallel~$\ceu{\AtOr}$, however, if one of the sides terminates, the whole composition
terminates and function~$\clear$ is used to properly finalize the aborted
side (rules \R{or-nop1} and \R{or-nop2}).
\begin{gather*}
  \<\ceu{{\Nop}\AtAnd{q}},n,\nil>\nst\<q,n,\nil>\Rtag{and-nop1}\\[2\jot]
  %%
  \AxiomC{$\isblocked(p,n)$}
  \UnaryInfC{$\<\ceu{p\AtAnd{\Nop}},n,\nil>\nst\<p,n,\nil>$}
  \DisplayProof
  \Rtag{and-nop2}\\[2\jot]
  %%
  \<\ceu{{\Nop}\AtOr{q}},n,\nil>\nst\<\clear(q),n,\nil>\Rtag{or-nop1}\\[2\jot]
  %%
  \AxiomC{$\isblocked(p,n)$}
  \UnaryInfC{$\<\ceu{p\AtOr{\Nop}},n,\nil>\nst\<\clear(p),n,\nil>$}
  \DisplayProof
  \Rtag{or-nop2}
\end{gather*}

@@

Finally, a~$\ceu{\Break}$ in one of the sides in parallel escapes the closest
enclosing~$\ceu{\Loop}$, properly aborting the other side with the~$\clear$
function:
\begin{gather*}
  \hskip-.5em
  \<\ceu{{\Break}\AtAnd{q}},n,\nil>\nst\<\ceu{\clear(q);\Break},n,\nil>
  \Rtag{and-brk1}\\[2\jot]
  %%
  \hskip-.5em
  \AxiomC{$\isblocked(p,n)$}
  \UnaryInfC{$\<\ceu{p\AtAnd{\Break}},n,\nil>
    \nst\<\ceu{\clear(p);\Break},n,\nil>$}
  \DisplayProof
  \Rtag{and-brk2}\\[2\jot]
  %%
  \<\ceu{{\Break}\AtOr{q}},n,\nil>\nst\<\ceu{\clear(q);\Break},n,\nil>
  \Rtag{or-brk1}\\[2\jot]
  %%
  \AxiomC{$\isblocked(p,n)$}
  \UnaryInfC{$\<\ceu{p\AtOr{\Break}},n,\nil>
    \nst\<\ceu{\clear(p);\Break},n,\nil>$}
  \DisplayProof
  \Rtag{or-brk2}
\end{gather*}

@@

In this section, we prove that a sufficiently long sequence of applications
of~$\trans$, and consequently~$\out$ and~$\nst$, will eventually lead to an
irreducible description, viz., one that cannot be modified by further
transitions.

>>>>

% \newcommand{\NST}{\1\xrightarrow[\mathit{nst}]\1}
% \newcommand{\OUT}{\1\xrightarrow[\mathit{out}]\1}
% \newcommand{\LL}{\langle}
% \newcommand{\RR}{\rangle}
% \newcommand{\DS}{\displaystyle}

% \newcommand{\1}{\;}
% \newcommand{\2}{\;\;}
% \newcommand{\3}{\;\;\;}
% \newcommand{\5}{\;\;\;\;\;}

\def\JOT{1.5\jot}

@@

The grammar below defines the syntax of a subset of \CEU that is
sufficient to describe all semantic peculiarities of the language.
\bgroup
\def\lbl#1{\quad\text{\emph{#1}}}%
\newdimen\X
\X=-1\jot
\begin{alignat*}{2}
  P\Coloneqq
      &\enspace\ceu{\Mem(\Id)}
      &&\lbl{any memory access to~``$\ceu\Id$''}\\[\X]
      %%
  \mid&\enspace\ceu{\AwaitExt(\Id)}
      &&\lbl{await external event~``$\ceu\Id$''}\\[\X]
      %%
  \mid&\enspace\ceu{\AwaitInt(\Id)}
      &&\lbl{await internal event~``$\ceu\Id$''}\\[\X]
      %%
  \mid&\enspace\ceu{\EmitInt(\Id)}
      &&\lbl{emit internal event~``$\ceu\Id$''}\\[\X]
      %%
  \mid&\enspace\ceu{\Break}
      &&\lbl{loop escape}\\[\X]
      %%
  \mid&\enspace\ceu{\IfElse{\Mem(\Id)}{P_1}{P_2}}
      &&\lbl{conditional}\\[\X]
      %%
  \mid&\enspace\ceu{P_1\,;\,P_2}
      &&\lbl{sequence}\\[\X]
      %%
  \mid&\enspace\ceu{\Loop P_1}
      &&\lbl{repetition}\\[\X]
      %%
  \mid&\enspace\ceu{\Every{\Id}\ P_1}
      &&\lbl{event iteration}\\[\X]
      %%
  \mid&\enspace\ceu{P_1\And P_2}
      &&\lbl{par/and}\\[\X]
      %%
  \mid&\enspace\ceu{P_2\Or P_2}
      &&\lbl{par/or}\\[\X]
      %%
  \mid&\enspace\ceu{\Fin P}
      &&\lbl{finalization}\\[\X]
      %%
  \mid&\enspace\ceu{P_1\AtLoop P_2}
      &&\lbl{unwinded loop}\\[\X]
      %%
  \mid&\enspace\ceu{P_1\AtAnd\ P_2}
      &&\lbl{unwinded par/and}\\[\X]
      %%
  \mid&\enspace\ceu{P_1\AtOr\ P_2}
      &&\lbl{unwinded par/or}\\[\X]
      %%
  \mid&\enspace\ceu{\CanRun(n)}
      &&\lbl{can run on stack level~$n$}\\[\X]
      %%
  \mid&\enspace\ceu{\Nop}
      &&\lbl{terminated program}
\end{alignat*}
\egroup

@@

Most expressions in the abstract language are mapped to their counterparts in
the concrete language.
The exceptions are the finalization block~$\ceu{\Fin{P}}$ and the
\texttt{@}-expressions which result from expansions in the transition rules to
be presented.

@@

The core of our semantics describes how a program reacts to a single external
input event, i.e., starting from an input event, how the program behaves and
becomes idle again to proceed to a subsequent reaction.
%
We use a set of small-step operational rules, which are designed in such a way
that at most one transition is possible at any time, resulting in deterministic
reactions.
%
The transition rules map a triple with a program~$p$, a stack level~$n$, and an
emitted event~$e$ to a modified triple as follows:
\[
  \<p,n,e>\trans\<p',n',e'>\,,
\]
where~$p,p'\in\P$ are abstract-language programs, $n,n'\in\N$ are non-negative
integers representing the current stack level, and~$e,e'\in\E\cup\{\nil\}$ are
the events emitted before and after the transition (both being possibly the
empty event~$\nil$).
$\E$ is a set of primitive events.

@@

The~$\out$ rules \R{push} and \R{pop} are non-recursive definitions which only
apply to the program as a whole, and are the only to manipulate the stack
level:
\begin{align*}
  &\AxiomC{$e\ne\nil$}
  \UnaryInfC{$\<p,n,e>\out\<\bcast(p,e),n+1,\nil>$}
  \DisplayProof
  \Rtag{push}\\[\JOT]
  %%
  &\AxiomC{$n>0$}
  \AxiomC{$\ceu{p=\Nop}\vee\isblocked(p,n)$}
  \BinaryInfC{$\<p,n,\nil>\out\<p,n-1,\nil>$}
  \DisplayProof
  \Rtag{pop}
\end{align*}

@@

The~$\nst$ rules are recursive definitions of the form
\[
\<p,n,\nil>\nst\<p',n,e>.
\]
%
%-
% \begin{align*}
% \LL p, n,\epsilon \RR &\NST
% \LL p',n,e        \RR
%     & \textbf{(rule-inner)}
% \end{align*}
%-
%
Nested transitions do not affect the stack level and never have an emitted
event as a precondition.  The distinction between~$\out$ and~$\nst$ prevents
rules \R{push} and \R{pop} from matching and, consequently, from
inadvertently modifying the current stack level before the nested reaction
is complete.

A complete reaction consists of a series of transitions of the form
\begin{align*}
  \<p,0,e_\ext>\outpush\<p_1,1,\nil>
  \Big[\null\nst[*]\null\out\null\Big]\!\!\ast
  \null\nst[*]\null\outpop\<p',0,\nil>\,.
\end{align*}
%
%-
% \begin{align*}
% a) &\5\5
%     \LL p,0,ext \RR
%         \1\xrightarrow[out]{push}\1
%     \LL q,1,\epsilon \RR
% \\
% b) &\5\5 \1[ \1\xrightarrow[in]{*}\1
%     \LL r,i,e \RR
%         \1\xrightarrow[out]\1
%     \LL s,j,\epsilon \RR \1]*
% \\
% c) &\5\5 \1\xrightarrow[in]{*}\1
%     \LL t,k,\epsilon \RR
%         \1\xrightarrow[out]{pop}\1
%     \LL u,0,\epsilon \RR
% \end{align*}
%-
%
First, a~$\outpush$ starts a nested reaction at level~1.
Then, a series of alternations between zero or more~$\nst$ transitions (nested reactions) and a
single~$\out$ transition (stack operation) takes place.
Finally, a last~$\outpop$ transition decrements the
stack level to~0 and terminates the reaction.

The~$\nst$ rules for atoms are defined as follows:
\begin{align*}
  \<\ceu{\Mem(\Id)},n,\nil>
  &\nst\<\ceu{\Nop},n,\nil>\Rtag{mem}\\[\JOT]
  %%
  \<\ceu{\EmitInt(\Id)},n,\nil>
  &\nst\<\ceu{\CanRun(n)},n,\ceu{\Id}>\Rtag{emit-int}\\[\JOT]
  %%
  \<\ceu{\CanRun(n)},n,\nil>
  &\nst\<\ceu{\Nop},n,\nil>\Rtag{can-run}
\end{align*}

@@

The rules for conditionals and sequences are straightforward:
%
\begin{gather*}
  \AxiomC{$\eval(\ceu{\Mem(\Id)})$}
  \UnaryInfC{$\<\ceu{\IfElse{\Mem(\Id)}{p}{q}},n,\nil>\nst\<p,n,\nil>$}
  \DisplayProof
  \Rtag{if-true}\\[\JOT]
  %%
  \AxiomC{$\lnot\eval(\ceu{\Mem(\Id)})$}
  \UnaryInfC{$\<\ceu{\IfElse{\Mem(\Id)}{p}{q}},n,\nil>\nst\<q,n,\nil>$}
  \DisplayProof
  \Rtag{if-false}\\[\JOT]
  %%
  \AxiomC{$\<p,n,\nil>\nst\<p',n,e>$}
  \UnaryInfC{$\<\ceu{p\,;\,q},n,\nil>\nst\<\ceu{p';\,q},n,e>$}
  \DisplayProof
  \Rtag{seq-adv}
\end{gather*}
\vskip-\belowdisplayskip
\vskip-\abovedisplayskip
\vskip\JOT
\begin{align*}
  \<\ceu{\Nop;\,q},n,\nil>&\nst\<q,n,\nil>\Rtag{seq-nop}\\[\JOT]
  %%
  \<\ceu{\Break;\,q},n,\nil>&\nst\<\ceu{\Break},n,\nil>\Rtag{seq-brk}
\end{align*}

@@

The rules for loops are similar to those for sequences, but use ``\code{@}''
as separators to bind breaks to their enclosing loops:
\begin{align*}
  \<\ceu{\Loop{p}},n,\nil>
  &\nst\<\ceu{p\AtLoop{p}},n,\nil>\Rtag{loop-expd}\\[\JOT]
  %%
  &\hskip-6.35em
  \AxiomC{$\<q,n,\nil>\nst\<q',n,e>$}
  \UnaryInfC{$\<\ceu{q\AtLoop{p}},n,\nil>\nst\<\ceu{q'\AtLoop{p}},n,e>$}
  \DisplayProof
  \Rtag{loop-adv}\\[\JOT]
  %%
  \<\ceu{\Nop\AtLoop{p}},n,\nil>
  &\nst\<\ceu{\Loop{p}},n,\nil>\Rtag{loop-nop}\\[\JOT]
  %%
  \<\ceu{\Break\AtLoop{p}},n,\nil>
  &\nst\<\ceu{\Nop},n,\nil>\Rtag{loop-brk}
\end{align*}

@@

The rules for~$\ceu{\And}$ and~$\ceu{\Or}$ compositions make sure that their
left branches always transition before their right branches:
%%
\begin{gather*}
  \hskip-.9em
  \<\ceu{p\And{q}},n,\nil>
  \nst\<\ceu{p\AtAnd(\CanRun(n);q)},n,\nil>
  \Rtag{and-expd}\\[\JOT]
  %%
  \AxiomC{$\<p,n,\nil>\nst\<p',n,e>$}
  \UnaryInfC{$\<\ceu{p\AtAnd{q}},n,\nil>\nst\<\ceu{p'\AtAnd{q}},n,e>$}
  \DisplayProof
  \Rtag{and-adv1}\\[\JOT]
  %%
  \AxiomC{$\isblocked(p,n)$}
  \AxiomC{$\<q,n,\nil>\nst\<q',n,e>$}
  \BinaryInfC{$\<\ceu{p\AtAnd{q}},n,\nil>\nst\<\ceu{p\AtAnd{q'}},n,e>$}
  \DisplayProof
  \Rtag{and-adv2}
\end{gather*}

\begin{gather*}
  \<\ceu{p\Or{q}},n,\nil>
  \nst\<\ceu{p\AtOr(\CanRun(n);q)},n,\nil>
  \Rtag{or-expd}\\[\JOT]
  %%
  \AxiomC{$\<p,n,\nil>\nst\<p',n,e>$}
  \UnaryInfC{$\<\ceu{p\AtOr{q}},n,\nil>\nst\<\ceu{p'\AtOr{q}},n,e>$}
  \DisplayProof
  \Rtag{or-adv1}\\[\JOT]
  %%
  \AxiomC{$\isblocked(p,n)$}
  \AxiomC{$\<q,n,\nil>\nst\<q',n,e>$}
  \BinaryInfC{$\<\ceu{p\AtOr{q}},n,\nil>\nst\<\ceu{p\AtOr{q'}},n,e>$}
  \DisplayProof
  \Rtag{or-adv2}
\end{gather*}

@@

In a parallel~$\ceu{\AtAnd}$, if one of the sides terminates, the composition is
simply substituted by the other side (rules \R{and-nop1} and
\R{and-nop2} below).
%
In a parallel~$\ceu{\AtOr}$, however, if one of the sides terminates, the
whole composition
terminates and function~$\clear$ is used to properly finalize the aborted
side (rules \R{or-nop1} and \R{or-nop2}).
\begin{gather*}
  \<\ceu{{\Nop}\AtAnd{q}},n,\nil>\nst\<q,n,\nil>\Rtag{and-nop1}\\[\JOT]
  %%
  \AxiomC{$\isblocked(p,n)$}
  \UnaryInfC{$\<\ceu{p\AtAnd{\Nop}},n,\nil>\nst\<p,n,\nil>$}
  \DisplayProof
  \Rtag{and-nop2}\\[\JOT]
  %%
  \<\ceu{{\Nop}\AtOr{q}},n,\nil>\nst\<\clear(q),n,\nil>\Rtag{or-nop1}\\[\JOT]
  %%
  \AxiomC{$\isblocked(p,n)$}
  \UnaryInfC{$\<\ceu{p\AtOr{\Nop}},n,\nil>\nst\<\clear(p),n,\nil>$}
  \DisplayProof
  \Rtag{or-nop2}
\end{gather*}

@@

Finally, a~$\ceu{\Break}$ in one of the sides in parallel escapes the closest
enclosing~$\ceu{\Loop}$, properly aborting the other side with the~$\clear$
function:
\begin{gather*}
  \hskip-.5em
  \<\ceu{{\Break}\AtAnd{q}},n,\nil>\nst\<\ceu{\clear(q);\Break},n,\nil>
  \Rtag{and-brk1}\\[\JOT]
  %%
  \hskip-.5em
  \AxiomC{$\isblocked(p,n)$}
  \UnaryInfC{$\<\ceu{p\AtAnd{\Break}},n,\nil>
    \nst\<\ceu{\clear(p);\Break},n,\nil>$}
  \DisplayProof
  \Rtag{and-brk2}\\[\JOT]
  %%
  \<\ceu{{\Break}\AtOr{q}},n,\nil>\nst\<\ceu{\clear(q);\Break},n,\nil>
  \Rtag{or-brk1}\\[\JOT]
  %%
  \AxiomC{$\isblocked(p,n)$}
  \UnaryInfC{$\<\ceu{p\AtOr{\Break}},n,\nil>
    \nst\<\ceu{\clear(p);\Break},n,\nil>$}
  \DisplayProof
  \Rtag{or-brk2}
\end{gather*}

@@

We proceed to prove that a sufficiently long sequence of applications
of~$\trans$ will eventually lead to an irreducible description, i.e., one
that cannot be modified by further transitions.  Before that, however, we
need to introduce some definitions and basic results regarding~$\nst$
and~$\out$.
